---
title: "R Prog IP"
author: "Kevin Kilonzo"
date: "23/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R PROGRAMMING: UNIVARIATE AND BIVARIATE ANALYSIS
## 1- DEFINING THE QUESTION
  a) Specifying the question
  Performing EDA on the dataset provided
  
  b) Defining the metric for success
  To be able to perform analysis on the dataset provided
  
  c) Understanding the context
  The IP advertising dataset gives information on the ad clicks of the respondents and whether they click on the ads or not
  
  d) Recording the experimental design 
    1. Defining the question
    2. Instaling and loading the libraries
    3. Loading the dataset
    4. Cleaning the dataset
    5. Performing univariate analysis
    6. Performing bivariate analysis
  
  e) Data relevance
    * Daiy time spent on site - Time spent on the site in question daily
    * Age - Age of the respondents
    * Area income - Income of the area where the respondent is accessing the site
    * Daily Internet usage - Time spent using the internet by the respondent daily
    * Ad topic line - The ad displayed to the respondent
    * City - the city the respondent is viewing rthe ad
    * Male - Whether respondent is male or not (1-Male , 0-Female)
    * Country - Country where the respondent is viewing the ad
    * Timestamp - Timestamp of the moment of viewing the ad
    * Clicked on ad - Whether the respondent clicked on the ad or not (1-Clicked, 0-Not clicked)
    
## 2- INSTALLING AND LOADING THE LIBRARIES
```{r}
 # Installing and loading libraries
# install.packages("r package",repo="https://github.com/Rdatatable/data.table")
library("data.table")

# install.packages("tidyverse")
library("tidyverse")
# Loading the dataset
data <- fread("http://bit.ly/IPAdvertisingData")
```

## 3.LOAD AND CHECK THE DATASET
```{r}
# Preview of the dataset
head(data)
# Show dataframe
view(data)
# Check datatypes
str(data)
# Check rows and columns
dim(data)
```

The data has 1000 rows and 10 columns

## 4. CLEANING THE DATASET
```{r}
# Checking for null values
# Checking missing values in columns
colSums(is.na(data))

# Checking for duplicates
dupli <- data[duplicated(data),]
dupli

# Checking for outliers
nums <- select_if(data, is.numeric)
head(nums)
boxplot(nums$`Daily Time Spent on Site`)
boxplot(nums$Age)
boxplot(nums$`Area Income`)
boxplot(data$`Daily Internet Usage`)
```

The data has no missing data and has no duplicates
The other two remaining numerical columns (Male, Clicked on ad ) are binary so no outliers
Outliers found in area income are expected as they are on the lower side which is possible with people with low or no incomes 

## 5. UNIVARIATE ANALYSIS
```{r}
# Mode function for getting mode
getmode <- function(v)
{uniqv <- unique(v)
uniqv[which.max(tabulate(match(v,uniqv)))]}

# Daily time spent on site colummn
mean_daily <- mean(nums$`Daily Time Spent on Site`)
median_daily <- median(nums$`Daily Time Spent on Site`)
mode_daily <- getmode(nums$`Daily Time Spent on Site`)
quartile_daily <- quantile(nums$`Daily Time Spent on Site`)
range_daily <- range(nums$`Daily Time Spent on Site`)
std_dev_daily <- sd(nums$`Daily Time Spent on Site`)
hist(nums$`Daily Time Spent on Site`)
# Print outs
# mean_daily
mean_daily
# median_daily
median_daily
# mode_daily
mode_daily
# quartile_daily
quartile_daily
# range_daily
range_daily
# std_dev_daily
std_dev_daily


# Age column
mean_age <- mean(nums$Age)
median_age <- median(nums$Age)
mode_age <- getmode(nums$Age)
quartile_age <- quantile(nums$Age)
range_age <- range(nums$Age)
std_dev_age <- sd(nums$Age)
hist(nums$Age)
# Print outs
# mean_age
mean_age
# median_age
median_age
# mode_age
mode_age
# quartile_age
quartile_age
# range_age
range_age
# std_dev_age
std_dev_age


# Area income column
mean_income <- mean(nums$`Area Income`)
median_income <- median(nums$`Area Income`)
mode_income <- getmode(nums$`Area Income`)
quartile_income <- quantile(nums$`Area Income`)
range_income <- range(nums$`Area Income`)
std_dev_income <- sd(nums$`Area Income`)
hist(nums$`Area Income`)
# Print outs
# mean_income
mean_income
# median_income
median_income
# mode_income
mode_income
# quartile_income
quartile_income
# range_income
range_income
# std_dev_income
std_dev_income



# Daily internet usage column
mean_usage <- mean(nums$`Daily Internet Usage`)
median_usage <- median(nums$`Daily Internet Usage`)
mode_usage <- getmode(nums$`Daily Internet Usage`)
quartile_usage <- quantile(nums$`Daily Internet Usage`)
range_usage <- range(nums$`Daily Internet Usage`)
std_dev_usage <- sd(nums$`Daily Internet Usage`)
hist(nums$`Daily Internet Usage`)
# Print outs
# mean_usage
mean_usage
# median_usage
median_usage
# mode_usage
mode_usage
# quartile_usage
quartile_usage
# range_usage
range_usage
# std_dev_usage
std_dev_usage

# Bar graphs for the binary columns 
# Assigning cols to variables
click <- nums$`Clicked on Ad`
gender <- nums$Male

# Frequency tables
freq_click <- table(click)
freq_gender <- table(gender)

# Plotting bar graphs
# Gender bargraph
barplot(freq_gender)
# Ad clicks bar graph
barplot(freq_click)

# Categorical columns ('ad topic line' , city, country)
ad_line <- data$`Ad Topic Line`
city <- data$City
country <- data$Country
# Frequency tables
freq_ad <- table(ad_line)
freq_city <- table(city)
freq_country <- table(country)

# Getting top 10 values and bottom 10 values
sf <- sort(freq_ad , decreasing = T)
scity <- sort(freq_city, decreasing = T)
scountry <- sort(freq_country, decreasing = T)
head(sf, n= 10)
head(scity, n= 10)
head(scountry, n= 10)
tail(scity, n= 10)
tail(scountry, n= 10)
```
* Daily time spent mean = 65.0002, median = 68.215 , mode = 68.215 , IQR (0% =32.6, 25%=51.36, 50%=68.215, 75%=78.5475, 100%=91.43)
- range 32.60 - 91.43, std dev = 15.85361
* Age mean = 36 ,median = 35, mode = 31, IQR (0% =19, 25%= 29, 50%= 35, 75%= 42, 100%= 61), range 19-61, std dev = 8.7855
* Area income mean = 55000, median = 57012.3, mode= 61833.9, IQR(0%=13996.50, 25%= 47031.80, 50%= 57012.30, 75%= 65470.64, 100% =79484.80), range 13996.5 - 79484.8, std dev = 13414.63
* Daily internet usage mean= 180.0001, median = 183.13, mode= 167.22, IQR(0%= 104.78, 25%= 138.83, 50%= 183.13, 75%= 218.7925, 100%= 269.96), range 104.78 - 269.96, std dev = 43.90243
* The distribution for daily time spent is bimodal and non normal
* The distribution for age is skewed to the left but normal
* The distribution for area income is skewed to the right but normal
* The distribution of daily internet usage is bimodal and not normal
* Most people spent 75 - 80 on the site
* Most people are between the age of 30 and 35
* Most people earn between 60000 and 65000 in income
* Most people spent 120-140 minutes on the internet
* The most common city was Lisamouth and Williamsport
* The most common countries were Czech Republic and France
* The are slightly more female than male respondents
* The respondents who clicked and those who did not click are equal
* The advertising line was unique to each respondent

## 6. BIVARIATE ANALYSIS
```{r}
# Correlation matrix of the numerical colummns
cor(nums)
library("corrplot")

corrplot(cor(nums), method = "number")
```

* Daily time spent on site has a positive correlation with Daily Internet usage and strong negative correlation with Clicked on ad
* Age has a positive correlation with clicked on ad
* Area income has a negative correlation with clicked on ad
* Daily internet usage has a positive correlation with Daily time spent on site and strong negative correlation with clicked on ad
* Clicked on ad has positive correlation with age, strong negative correlation with Daily time spent on site and Daily internet usage and negative correlation with area income

